
a cleaner version of countyouts.rmd, for replicatability

##loading packages
```{r}
library(tidyverse)
library(sf)
```

##read data
```{r}
countyoutflow <- read.csv("~/Desktop/elliesacbeeR/data/countyoutflow2021.csv")

us <- st_read("/Users/elin/Downloads/cb_2018_us_county_500k")
```
##filter it down
```{r}
california <- countyoutflow %>% group_by(y2_countyname) %>% filter(y1_statefips == "6" & y2_statefips !=96 & y2_statefips !=97 & y2_statefips !=98)
```

##adding the full fips code
str_pad makes the fips codes into 3 digits (county) and 2 digits (state) so that they can be concated together into a real ~~boy~~ fips code
```{r}
california$y2_countyfips <- str_pad(california$y2_countyfips, 3, pad = 0)

california$y2_statefips <- str_pad(california$y2_statefips, 2, pad = 0)

california <- california %>% mutate(state2fullfips = str_c(y2_statefips, y2_countyfips))
```

### adding the full fips code to the us shapefile
```{r}
us <- us %>% mutate(fullfips = str_c(STATEFP, COUNTYFP))
```

##grouping
going to group by y2 county names so that i can 
```{r}
california %>% group_by(y2_countyname, state2fullfips) %>% summarise(sum(n2))
```
### integrity check: hennepin county
```{r}
california %>% group_by(y2_countyname) %>% filter(y2_countyname == "Hennepin County") %>% summarise(sum(n2))
```
it looks right to me but will definitely need a second set of eyes

```{r}
simplified <- california %>% group_by(y2_countyname, state2fullfips) %>% summarise(sum(n2))
```

##merging
```{r}
uswithdata <- merge(simplified, us, by.x = "state2fullfips", by.y = "fullfips")
```

